Latency-tolerant runtimes for large graph computations

### 2011-01-03 doc/hotpar2011/softxmt-hotpar2011.tex
outline

\section{Graph problems}

\section{Multithreading on commodity processors}

\subsection{Memory concurrency}
\subsection{Synchronization}

\section{Results}

\section{Future work}

>>>>

sparse graph problems important

perform badly on commodity processors, since most accesses are misses

context-swtiching for latency tolerance has been successful with MTA, but single thread performance sucks

we want latency tolerance on commodity processors for best of both worlds

could just wait for wider multicores, but if context switching works, we can get enough concurrency on fewer cores and shut off the rest.

\section{Programming model for graph problems}

many many threads

prefetch and switch on

full/empty bits

\subsection{Opportunities}

little FP use ==> don't worry about FP registers

would like enough concurrency to keep pipeline filled, but commodity CPUs don't have enough external bandwidth. So, get enough to saturate bandwidth.

\section{Approach}

Describe general approach.

\subsection{Low-latency context switch}

switch only stack pointer and program counter; depend on compiler to save/restore anything else. in general case, hopefully this will be fast.

\subsection{Memory concurrency}

Commodity procs don't provide enough. use software? FPGA?

\subsection{Synchronization}

use existing atomic instructions? Use FPGA listening snoops on QPI?

\section{Current Implementation and Results}

current: just context switch on single node

linked lists:

there is some room for memory concurrency

thread contexts don't take up too much space in cache

sparse matmul:

speedup?

\section{Related work}

hardware:

XMT/MTA

cyclops

niagra

software:

qthreads

cilk? tbb?
